---
title: "CSU ESS330 Lab 10"
author:
  name: "Olivia Gilpin"
  email: "ogilpin@colostate.edu"
date: "4-30-2025"  
project: 
  output-dir: docs
  type: website
format: 
  html: 
    self-contained: true
    toc: true
    toc-depth: 6
    toc-location: left 
    toc-title: "Table of Contents"
execute:
  echo: true
editor: visual
---

# Lab 10: Distances and Projections
## Distance from the Border Zone

### Libraries
```{r}
# spatial data science
library(tidyverse)
library(sf)
library(units)

# Data
library(sf)

# Visualization
library(gghighlight)
library(ggrepel)
library(knitr)
```
#### Question 1
##### 1.1 Define a Projection
```{r}
eqdc <- '+proj=eqdc +lat_0=40 +lon_0=-96 +lat_1=20 +lat_2=60 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs'
```
##### 1.2 - Get USA state boundaries
```{r}
#| echo: true
remotes::install_github("ropensci/USAboundaries")
remotes::install_github("ropensci/USAboundariesdata")
```
```{r}
#| echo: true 
remotes::install_github("ropenscilabs/rnaturalearthdata")
```
##### 1.3 - Get country boundaries for Mexico, the United States of America, and Canada
```{r}
eqdc <- st_crs(5070) 

states <- USAboundaries::us_states(resolution = "low") |>
  filter(!state_abbr %in% c("HI", "AK", "PR")) |>
  st_transform(eqdc)

countries <- rnaturalearthdata::countries110 |>
  st_transform(eqdc)
```


##### 1.4 - Get city locations from the CSV file
```{r}
library(readr)
us_cities <- readr::read_csv("data/uscities.csv")

head(us_cities)  
str(us_cities)   
summary(us_cities)  
```
```{r}
df <- data.frame(name = state.name, 
                X = state.center$x, 
                Y = state.center$y)
head(df)
```
```{r}
(df_sf_gcs = st_as_sf(df, 
                      coords = c("X", "Y"), 
                      crs = 4269))
```
```{r}
ggplot() + 
  geom_sf(data = df_sf_gcs) + 
  coord_sf(datum = st_crs(df_sf_gcs)) +
  theme_linedraw()
```
```{r}
(df_sf_pcs = st_transform(df_sf_gcs, 5070))
```
```{r}
ggplot() + 
  geom_sf(data = df_sf_pcs) + 
  coord_sf(datum = st_crs(df_sf_pcs)) +
  theme_linedraw()
```
```{r}
library(readr)
cities <- read_csv("data/uscities.csv", show_col_types = FALSE) |>
  st_as_sf(coords = c("lng", "lat"), crs = 4326) |>
  st_transform(eqdc)
```

#### Question 2
##### 2.1 - Distance to USA Border (coastline or national) (km)
```{r}
usa_border <- states |>
  st_union() |> 
  st_cast("MULTILINESTRING")

cities <- cities |>
  mutate(
    dist_to_usa_border = st_distance(cities, usa_border),
    dist_to_usa_border = set_units(dist_to_usa_border, "km"),
    dist_to_usa_border = drop_units(dist_to_usa_border)
  )

farthest_from_usa <- cities |>
  select(city, state_name, dist_to_usa_border) |>
  arrange(desc(dist_to_usa_border)) |>
  slice_head(n = 5)

library(flextable)
flextable(farthest_from_usa |> 
            st_drop_geometry() |>
            select(city, state_name, dist_to_usa_border)) |>
  set_header_labels(
    city = "City",
    state_name = "State",
    dist_to_usa_border = "Distance (km)"
  ) |>
  colformat_double(j = "dist_to_usa_border", digits = 2) |>
  autofit()
```

##### 2.2 - Distance to States (km)
```{r}
state_borders <- states |>
  st_combine() |>  
  st_cast("MULTILINESTRING")

cities <- cities |>
  mutate(
    dist_to_state_border = st_distance(cities, state_borders),
    dist_to_state_border = set_units(dist_to_state_border, "km"),
    dist_to_state_border = drop_units(dist_to_state_border)
  )

farthest_from_state <- cities |>
  select(city, state_name, dist_to_state_border) |>
  arrange(desc(dist_to_state_border)) |>
  slice_head(n = 5)

flextable(farthest_from_state |> 
            st_drop_geometry() |>
            select(city, state_name, dist_to_state_border)) |>
  set_header_labels(
    city = "City",
    state_name = "State",
    dist_to_state_border = "Distance (km)"
  ) |>
  colformat_double(j = "dist_to_state_border", digits = 2) |>
  autofit()
```

##### 2.3 - Distance to Mexico (km)
```{r}
mexico <- countries |>
  filter(admin == "Mexico") |>
  st_cast("MULTILINESTRING")

cities <- cities |>
  mutate(
    dist_to_mexico = st_distance(cities, mexico),
    dist_to_mexico = set_units(dist_to_mexico, "km"),
    dist_to_mexico = drop_units(dist_to_mexico)
  )

farthest_from_mexico <- cities |>
  select(city, state_name, dist_to_mexico) |>
  arrange(desc(dist_to_mexico)) |>
  slice_head(n = 5)

flextable(farthest_from_mexico |> 
            st_drop_geometry() |>
            select(city, state_name, dist_to_mexico)) |>
  set_header_labels(
    city = "City",
    state_name = "State",
    dist_to_mexico = "Distance (km)"
  ) |>
  colformat_double(j = "dist_to_mexico", digits = 2) |>
  autofit()
```

##### 2.4 - Distance to Canada (km)
```{r}
canada <- countries |>
  filter(admin == "Canada") |>
  st_cast("MULTILINESTRING")

cities <- cities |>
  mutate(
    dist_to_canada = st_distance(cities, canada),
    dist_to_canada = set_units(dist_to_canada, "km"),
    dist_to_canada = drop_units(dist_to_canada)
  )

farthest_from_canada <- cities |>
  select(city, state_name, dist_to_canada) |>
  arrange(desc(dist_to_canada)) |>
  slice_head(n = 5)

flextable(farthest_from_canada |> 
            st_drop_geometry() |>
            select(city, state_name, dist_to_canada)) |>
  set_header_labels(
    city = "City",
    state_name = "State",
    dist_to_canada = "Distance (km)"
  ) |>
  colformat_double(j = "dist_to_canada", digits = 2) |>
  autofit()
```

#### Question 3
##### 3.1 Data
```{r}
library(ggplot2)
library(sf)  
library(dplyr)  
library(rnaturalearth)
countries <- ne_countries()

names(countries)

na <- filter(countries, name %in% c("Mexico", "Canada", "United States of America")) |>
  st_as_sf()
big_cities <- cities |>
  slice_max(population, n = 10)

ggplot() +
  geom_sf(data = na, fill = "grey") +
  geom_sf(data = states, fill = "white", lty = 2, size = 0.5) + 
  geom_sf(data = big_cities, color = "red", size = .5) + 
  ggrepel::geom_label_repel(data = big_cities,
                            aes(geometry = geometry, label = city, label.size = .1), 
                            stat = "sf_coordinates") + theme_void()
```

##### 3.2 City Distance from the Border
```{r}
library(ggplot2)
library(sf)
library(dplyr)
library(ggrepel)

ggplot() +
  geom_sf(data = states, fill = "lightgray", color = "white", size = 0.2) +
  geom_sf(data = cities, aes(color = dist_to_usa_border), alpha = 0.5, size = 0.5) +
  scale_color_viridis_c(name = "Distance (km)", option = "plasma") +
  geom_sf(data = farthest_from_usa, color = "red", size = 2) +
  geom_label_repel(data = farthest_from_usa,
                   aes(geometry = geometry, label = city),
                   stat = "sf_coordinates", 
                   size = 3,
                   box.padding = 0.5,
                   point.padding = 0.3) +
  theme_minimal() +
  labs(title = "Distance of US Cities from National Border",
       subtitle = "Red points indicate the 5 cities farthest from the border",
       caption = "Data source: US Cities Database") +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "right")
```

##### 3.3 City Distance from Nearest State
```{r}
ggplot() +
  geom_sf(data = states, fill = "lightgray", color = "white", size = 0.2) +
  geom_sf(data = cities, aes(color = dist_to_state_border), alpha = 0.5, size = 0.5) +
  scale_color_viridis_c(name = "Distance (km)", option = "viridis") +
  geom_sf(data = farthest_from_state, color = "red", size = 2) +
  geom_label_repel(data = farthest_from_state,
                   aes(geometry = geometry, label = city),
                   stat = "sf_coordinates", 
                   size = 3,
                   box.padding = 0.5,
                   point.padding = 0.3) +
  theme_minimal() +
  labs(title = "Distance of US Cities from Nearest State Border",
       subtitle = "Red points indicate the 5 cities farthest from any state border",
       caption = "Data source: US Cities Database") +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "right")
```

##### 3.4 Equidistance boundary from Mexico and Canada
```{r}
library(gghighlight)

cities <- cities %>%
  mutate(dist_diff = abs(dist_to_mexico - dist_to_canada))

equidistant_cities <- cities %>%
  filter(dist_diff <= 100)

top_equidistant_cities <- equidistant_cities %>%
  slice_max(population, n = 5)

ggplot() +
  geom_sf(data = states, fill = "lightgray", color = "white", size = 0.2) +
  geom_sf(data = cities, aes(color = dist_diff), alpha = 0.7, size = 0.5) +
  scale_color_viridis_c(
    name = "Difference in Distance (km)", 
    option = "magma", 
    direction = -1,
    breaks = c(0, 100, 500, 1000, 1500),
    labels = c("0 (Equidistant)", "100", "500", "1000", "1500+")
  ) +
  geom_sf(data = equidistant_cities, color = "cyan", alpha = 0.8, size = 0.8, shape = 18) +
  geom_sf(data = top_equidistant_cities, color = "red", size = 3, shape = 15) +
  geom_label_repel(data = top_equidistant_cities,
                   aes(geometry = geometry, label = paste0(city, "\nPop: ", format(population, big.mark = ","))),
                   stat = "sf_coordinates", 
                   size = 3,
                   box.padding = 0.5,
                   point.padding = 0.3,
                   force = 5) +
  geom_sf(data = countries %>% filter(admin == "Mexico"), fill = NA, color = "orange", size = 1) +
  geom_sf(data = countries %>% filter(admin == "Canada"), fill = NA, color = "blue", size = 1) +
  theme_minimal() +
  labs(title = "Cities Equidistant from Mexican and Canadian Borders (±100 km)",
       subtitle = "Red squares indicate the 5 most populous cities in the equidistant zone",
       caption = "Data source: US Cities Database\nMexican border in orange, Canadian border in blue") +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "right",
        panel.grid.major = element_line(color = gray(0.9)),
        panel.grid.minor = element_blank())
```

#### Question 4
##### 4.1 Quantifing Border Zone
```{r}
border_zone_cities <- cities %>%
  filter(dist_to_usa_border <= 160) 

num_border_zone_cities <- nrow(border_zone_cities)

border_zone_population <- sum(border_zone_cities$population)

total_population <- sum(cities$population)

percentage_in_border_zone <- (border_zone_population / total_population) * 100

border_zone_summary <- data.frame(
  Metric = c("Number of Cities in Border Zone", 
             "Population in Border Zone", 
             "Percentage of Total Population in Border Zone"),
  Value = c(num_border_zone_cities,
            border_zone_population,
            paste0(round(percentage_in_border_zone, 2), "%"))
)

flextable(border_zone_summary) %>%
  autofit()
```

##### 4.2 Mapping Border Zone
```{r}
ggplot() +
  geom_sf(data = states, fill = "lightgray", color = "white", size = 0.2) +
  geom_sf(data = cities, color = "gray", alpha = 0.2, size = 0.1) +
  geom_sf(data = border_zone_cities, 
          aes(color = dist_to_usa_border), 
          alpha = 0.7, 
          size = 0.5) +
  scale_color_gradient(
    name = "Distance to Border (km)",
    low = "orange",
    high = "darkred",
    limits = c(0, 160)
  ) +
  geom_sf(data = border_zone_cities %>% slice_max(population, n = 10), 
          color = "blue", 
          size = 2) +
  geom_label_repel(
    data = border_zone_cities %>% slice_max(population, n = 10),
    aes(geometry = geometry, label = city),
    stat = "sf_coordinates",
    size = 3,
    box.padding = 0.5,
    point.padding = 0.3,
    force = 2
  ) +
  theme_minimal() +
  labs(
    title = "U.S. Cities Within 100 Miles of External Borders",
    subtitle = "Blue points represent the 10 most populous cities in the Border Zone",
    caption = "Data source: US Cities Database"
  ) +
  theme(plot.title = element_text(face = "bold"))
```

##### 4.3 : Instead of labeling the 10 most populous cites, label the most populous city in each state within the Danger Zone.
```{r}
states_all <- USAboundaries::us_states(resolution = "low") |>
  filter(!state_abbr %in% c("PR")) |>  
  st_transform(eqdc)

usa_border_all <- states_all |>
  st_union() |> 
  st_cast("MULTILINESTRING")

cities_all <- read_csv("data/uscities.csv", show_col_types = FALSE) |>
  st_as_sf(coords = c("lng", "lat"), crs = 4326) |>
  st_transform(eqdc) |>
  mutate(
    dist_to_usa_border = st_distance(geometry, usa_border_all),
    dist_to_usa_border = set_units(dist_to_usa_border, "km"),
    dist_to_usa_border = drop_units(dist_to_usa_border)
  )

border_zone_cities_all <- cities_all |>
  filter(dist_to_usa_border <= 160)

most_populous_by_state_all <- border_zone_cities_all |>
  group_by(state_name) |>
  slice_max(population, n = 1) |>
  ungroup()

library(cowplot)

main_map <- ggplot() +
  geom_sf(data = states_all |> filter(!state_abbr %in% c("HI", "AK")), 
          fill = "lightgray", color = "white", size = 0.2) +
  geom_sf(data = border_zone_cities_all |> filter(!state_name %in% c("Hawaii", "Alaska")), 
          aes(color = dist_to_usa_border), 
          alpha = 0.7, 
          size = 0.5) +
  scale_color_gradient(
    name = "Distance to Border (km)",
    low = "orange",
    high = "darkred",
    limits = c(0, 160)
  ) +
  geom_sf(data = most_populous_by_state_all |> filter(!state_name %in% c("Hawaii", "Alaska")), 
          color = "blue", 
          size = 2) +
  geom_label_repel(
    data = most_populous_by_state_all |> filter(!state_name %in% c("Hawaii", "Alaska")),
    aes(geometry = geometry, label = paste0(city, ", ", state_name)),
    stat = "sf_coordinates",
    size = 2.2,
    box.padding = 0.3,
    point.padding = 0.2,
    force = 5,
    max.overlaps = Inf,
    min.segment.length = 0.1,
    seed = 42
  ) +
  theme_minimal() +
  theme(legend.position = "right")

alaska_map <- ggplot() +
  geom_sf(data = states_all |> filter(state_abbr == "AK"), 
          fill = "lightgray", color = "white", size = 0.2) +
  geom_sf(data = border_zone_cities_all |> filter(state_name == "Alaska"), 
          aes(color = dist_to_usa_border), 
          alpha = 0.7, 
          size = 0.5) +
  scale_color_gradient(
    name = "Distance to Border (km)",
    low = "orange",
    high = "darkred",
    limits = c(0, 160)
  ) +
  geom_sf(data = most_populous_by_state_all |> filter(state_name == "Alaska"), 
          color = "blue", 
          size = 2) +
  geom_label_repel(
    data = most_populous_by_state_all |> filter(state_name == "Alaska"),
    aes(geometry = geometry, label = paste0(city, ", ", state_name)),
    stat = "sf_coordinates",
    size = 2.2,
    box.padding = 0.3,
    point.padding = 0.2
  ) +
  theme_void() +
  theme(legend.position = "none") +
  labs(title = "Alaska")

hawaii_map <- ggplot() +
  geom_sf(data = states_all |> filter(state_abbr == "HI"), 
          fill = "lightgray", color = "white", size = 0.2) +
  geom_sf(data = border_zone_cities_all |> filter(state_name == "Hawaii"), 
          aes(color = dist_to_usa_border), 
          alpha = 0.7, 
          size = 0.5) +
  scale_color_gradient(
    name = "Distance to Border (km)",
    low = "orange",
    high = "darkred",
    limits = c(0, 160)
  ) +
  geom_sf(data = most_populous_by_state_all |> filter(state_name == "Hawaii"), 
          color = "blue", 
          size = 2) +
  geom_label_repel(
    data = most_populous_by_state_all |> filter(state_name == "Hawaii"),
    aes(geometry = geometry, label = paste0(city, ", ", state_name)),
    stat = "sf_coordinates",
    size = 2.2,
    box.padding = 0.3,
    point.padding = 0.2
  ) +
  theme_void() +
  theme(legend.position = "none") +
  labs(title = "Hawaii")

final_map <- ggdraw() +
  draw_plot(main_map) +
  draw_plot(alaska_map, x = 0.15, y = 0.1, width = 0.2, height = 0.2) +
  draw_plot(hawaii_map, x = 0.5, y = 0.1, width = 0.15, height = 0.15) +
  draw_label("U.S. Cities Within 100 Miles of External Borders", 
             x = 0.5, y = 0.98, size = 16, fontface = "bold", hjust = 0.5) +
  draw_label("Blue points represent the most populous city in each state within the Border Zone", 
             x = 0.5, y = 0.95, size = 12, hjust = 0.5) +
  draw_label("Data source: US Cities Database", 
             x = 0.85, y = 0.02, size = 8, color = "gray30")

print(final_map)

border_zone_stats <- data.frame(
  Metric = c("Number of States with Cities in Border Zone", 
             "Number of Cities in Border Zone", 
             "Population in Border Zone", 
             "Percentage of Total Population"),
  Value = c(n_distinct(border_zone_cities_all$state_name),
            nrow(border_zone_cities_all),
            format(sum(border_zone_cities_all$population, na.rm = TRUE), big.mark = ","),
            paste0(round((sum(border_zone_cities_all$population, na.rm = TRUE) / 
                          sum(cities_all$population, na.rm = TRUE)) * 100, 2), "%"))
)

flextable(border_zone_stats) %>%
  autofit()
```

###### Refined 4.3
```{r}
library(sf)
library(ggplot2)
library(dplyr)
library(readr)
library(ggrepel)
library(cowplot)
library(flextable)
library(units)

eqdc <- 5070 

states_all <- USAboundaries::us_states(resolution = "low") |>
  filter(!state_abbr %in% c("PR")) |>  
  st_transform(eqdc)

usa_border_all <- states_all |>
  st_union() |> 
  st_cast("MULTILINESTRING")

cities_all <- read_csv("data/uscities.csv", show_col_types = FALSE) |>
  st_as_sf(coords = c("lng", "lat"), crs = 4326) |>
  st_transform(eqdc) |>
  mutate(
    dist_to_usa_border = st_distance(geometry, usa_border_all),
    dist_to_usa_border = set_units(dist_to_usa_border, "km"),
    dist_to_usa_border = drop_units(dist_to_usa_border)
  )

border_zone_cities_all <- cities_all |>
  filter(dist_to_usa_border <= 160)

most_populous_by_state_all <- border_zone_cities_all |>
  group_by(state_name) |>
  slice_max(population, n = 1) |>
  ungroup()

main_map <- ggplot() +
  geom_sf(data = states_all |> filter(!state_abbr %in% c("HI", "AK")), 
          fill = "lightgray", color = "white", size = 0.2) +
  geom_sf(data = border_zone_cities_all |> filter(!state_name %in% c("Hawaii", "Alaska")), 
          aes(color = dist_to_usa_border), 
          alpha = 0.7, 
          size = 0.5) +
  scale_color_gradient(
    name = "Distance to Border (km)",
    low = "orange",
    high = "darkred",
    limits = c(0, 160)
  ) +
  geom_sf(data = most_populous_by_state_all |> filter(!state_name %in% c("Hawaii", "Alaska")), 
          color = "blue", 
          size = 2) +
  geom_label_repel(
    data = most_populous_by_state_all |> filter(!state_name %in% c("Hawaii", "Alaska")),
    aes(geometry = geometry, label = paste0(city, ", ", state_name)),
    stat = "sf_coordinates",
    size = 2.2,
    box.padding = 0.3,
    point.padding = 0.2,
    force = 5,
    max.overlaps = Inf,
    min.segment.length = 0.1,
    seed = 42,
    bg.color = "white",
    bg.r = 0.15
  ) +
  theme_minimal() +
  theme(legend.position = "right")

alaska_map <- ggplot() +
  geom_sf(data = states_all |> filter(state_abbr == "AK"), 
          fill = "lightgray", color = "white", size = 0.2) +
  geom_sf(data = border_zone_cities_all |> filter(state_name == "Alaska"), 
          aes(color = dist_to_usa_border), 
          alpha = 0.7, 
          size = 0.5) +
  scale_color_gradient(
    name = "Distance to Border (km)",
    low = "orange",
    high = "darkred",
    limits = c(0, 160)
  ) +
  geom_sf(data = most_populous_by_state_all |> filter(state_name == "Alaska"), 
          color = "blue", 
          size = 3) +
  geom_label_repel(
    data = most_populous_by_state_all |> filter(state_name == "Alaska"),
    aes(geometry = geometry, label = paste0(city, ", ", state_name)),
    stat = "sf_coordinates",
    size = 2.2,
    box.padding = 0.3,
    point.padding = 0.2,
    force = 5,
    seed = 42,
    bg.color = "white",
    bg.r = 0.15
  ) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.background = element_rect(fill = "white", color = "black", linewidth = 0.8),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  labs(title = "Alaska")

hawaii_map <- ggplot() +
  geom_sf(data = states_all |> filter(state_abbr == "HI"), 
          fill = "lightgray", color = "white", size = 0.2) +
  geom_sf(data = border_zone_cities_all |> filter(state_name == "Hawaii"), 
          aes(color = dist_to_usa_border), 
          alpha = 0.7, 
          size = 1) +
  scale_color_gradient(
    name = "Distance to Border (km)",
    low = "orange",
    high = "darkred",
    limits = c(0, 160)
  ) +
  geom_sf(data = most_populous_by_state_all |> filter(state_name == "Hawaii"), 
          color = "blue", 
          size = 6) +
  annotate("label", 
           x = 0, y = 0,  
           label = "Honolulu, Hawaii",
           color = "black",
           fill = "white",
           size = 3.5,
           fontface = "bold") +
  geom_label_repel(
    data = most_populous_by_state_all |> filter(state_name == "Hawaii"),
    aes(geometry = geometry, label = paste0(city, ", ", state_name)),
    stat = "sf_coordinates",
    size = 3.5,             
    box.padding = 1.0,      
    point.padding = 1.0,    
    segment.size = 1.0,     
    min.segment.length = 0, 
    force = 100,            
    max.overlaps = 100,     
    nudge_x = 2.0,          
    nudge_y = -2.0,         
    direction = "both",     
    seed = 9999,            
    bg.color = "white",     
    bg.r = 0.2,            
    fill = "white",         
    color = "black",        
    fontface = "bold"       
  ) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.background = element_rect(fill = "white", color = "black", linewidth = 0.8),
    plot.margin = margin(15, 15, 15, 15) 
  ) +
  labs(title = "Hawaii") +
  coord_sf(expand = TRUE, xlim = c(-162, -154), ylim = c(18, 23))

final_map <- ggdraw() +
  draw_plot(main_map) +
  draw_plot(alaska_map, x = 0.08, y = 0.08, width = 0.2, height = 0.2) +
  draw_plot(hawaii_map, x = 0.75, y = 0.08, width = 0.22, height = 0.3) + # Increased height
  draw_label("U.S. Cities Within 100 Miles of External Borders (Border Zone)", 
             x = 0.5, y = 0.98, size = 16, fontface = "bold", hjust = 0.5) +
  draw_label("Blue points represent the most populous city in each state within the Border Zone", 
             x = 0.5, y = 0.95, size = 12, hjust = 0.5) +
  draw_label("Note: Not all states have cities within the Border Zone", 
             x = 0.5, y = 0.92, size = 10, hjust = 0.5, color = "darkred") +
  draw_label("Data source: US Cities Database", 
             x = 0.85, y = 0.02, size = 8, color = "gray30")

print(final_map)

border_zone_stats <- data.frame(
  Metric = c("Number of States with Cities in Border Zone", 
             "Number of Cities in Border Zone", 
             "Population in Border Zone", 
             "Percentage of Total Population"),
  Value = c(n_distinct(border_zone_cities_all$state_name),
            nrow(border_zone_cities_all),
            format(sum(border_zone_cities_all$population, na.rm = TRUE), big.mark = ","),
            paste0(round((sum(border_zone_cities_all$population, na.rm = TRUE) / 
                          sum(cities_all$population, na.rm = TRUE)) * 100, 2), "%"))
)

flextable(border_zone_stats) %>%
  autofit()
```





